---
title: "EnergySplit"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(DT)

export <- read_csv("~/R/energy_split/data/ami_daily_2017-02-27.csv")
export$date <- as.Date(floor_date(ymd_hms(export$UTC_START_TIME,tz="America/New_York"),unit="days"))
export <- export %>% 
  select(-X1) %>% 
  filter(date > '2016-01-01')

weather <- read_csv("~/R/energy_split/data/recentdatedweather.csv", 
    col_types = cols(CDD = col_double(), 
        DryBulb = col_double(), HDD = col_double()))
weather$date <- as.Date(ymd(weather$WeatherDate,tz="America/New_York"))
weather <- weather %>% 
  select(-X1)

export <- inner_join(export,weather, c("date"="date"))

```

## Usage v. Temperature

Having a variable for month of year or amount of sunshine would be more ideal.

```{r chart,echo=FALSE}
ggplot(export,aes(DryBulb,USAGE))+geom_point()+geom_rug(alpha=.3)+geom_smooth()
```

## Determine Setpoint

The setpoint is determined by the temperature at the lowest value in the above chart. The assumed setpoint is:

```{r setpoint,echo=FALSE}
p<-ggplot(export,aes(DryBulb,USAGE)) + geom_smooth()
model_data <- ggplot_build(p)$data[[1]][,1:5] 
setpoint <- model_data[order(model_data$y),][[1]][1]
setpoint
```

## (Re)Calculate HDD & CDD

Then, run the ol' linear model.

```{r lm,echo=FALSE}
export$HDD = ifelse(export$DryBulb > setpoint, 0, setpoint - export$DryBulb)
export$CDD = ifelse(export$DryBulb < setpoint, 0, export$DryBulb - setpoint)
digits <- tidy(lm(USAGE ~ HDD + CDD, export))[[2]]
datatable(tidy(lm(USAGE ~ HDD + CDD, export)), options = list(dom = 't'))
datatable(glance(lm(USAGE ~ HDD + CDD, export)), options = list(dom = 't'))
```


## Now Normalize for Typical Meterological Year 3

Based on the above setpoint, TMY 3 has these HDDs and CDDs:

```{r HDDCDD,echo=FALSE}
TMY3 <- read_csv("~/R/energy_split/data/TMY3.CSV")

TMY3summary <- TMY3 %>% 
  mutate(HDH = ifelse(Fdg < setpoint,setpoint - Fdg,0),CDH = ifelse(Fdg > setpoint,Fdg - setpoint,0)) %>% 
  summarise(HDD = sum(HDH)/24,CDD = sum(CDH)/24)

TMY3summary
```

Which, given our fit above leads to the following breakout (yearly load on top, percentage of total load on the bottom):

```{r breakout,echo=FALSE}
usage <- data.frame(digits[1] * 365)
colnames(usage)[1] <- 'Baseload'
usage$HeatingLoad <- digits[2] * TMY3summary[[1]]
usage$CoolingLoad <- digits[3] * TMY3summary[[2]]
datatable(rbind(usage,usage/sum(usage[1:3])), options = list(dom = 't'))
```

## Usage Charts

Actual Daily Usage:

```{r Actual,echo=FALSE}

ggplot(export,aes(date,USAGE)) + geom_point(alpha=.1)  + geom_smooth(span=0.3)
export$Baseload <- rep(digits[1],nrow(export))
export %>% mutate(HeatingLoad = HDD * digits[2],CoolingLoad = CDD * digits[3]) %>% gather(LoadType,Load,12:14) %>% ggplot(aes(date,Load,color = LoadType)) + geom_point(alpha=.3) + geom_smooth(span=0.3) + theme(legend.position="bottom")

```

TMY 3 Usage:

```{r TMY32,echo=FALSE}
TMY3$hour <-( c(0, rep(1:(nrow(TMY3)-1))) + 1 )
TMY3$Baseload <- rep(digits[1]/24,nrow(TMY3))

TMY3 %>% mutate(HeatingLoad = digits[2]*ifelse(Fdg < setpoint,setpoint - Fdg,0)/24,CoolingLoad = digits[3]*ifelse(Fdg > setpoint,Fdg - setpoint,0)/24,HourlyLoad = Baseload + HeatingLoad + CoolingLoad) %>% ggplot(aes(hour,HourlyLoad)) + geom_point(alpha=.1) +geom_smooth()

TMY3 %>% mutate(HeatingLoad = digits[2]*ifelse(Fdg < setpoint,setpoint - Fdg,0)/24,CoolingLoad = digits[3]*ifelse(Fdg > setpoint,Fdg - setpoint,0)/24) %>%  gather(LoadType,Load,3:5) %>% ggplot(aes(hour,Load,color = LoadType)) + geom_smooth()  + geom_point(alpha=.1) + theme(legend.position="bottom")

```

## Actual Residuals

```{r residuals,echo=FALSE}

export %>% mutate(HeatingLoad = HDD * digits[2],CoolingLoad = CDD * digits[3],Residual = Baseload + HeatingLoad + CoolingLoad - USAGE)  %>% ggplot(aes(date,Residual,group = 1)) + geom_point()

```
