---
title: "EnergySplit"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(DT)

#pull in energy data
export <- read_csv("~/R/energy_split/data/ami_daily_2017-02-27.csv")
export$date <- as.Date(floor_date(ymd_hms(export$UTC_START_TIME,tz="America/New_York"),unit="days"))
export <- export %>% 
  select(-X1) %>% 
  filter(date > '2016-01-01') # in this case habits changed around then, but there are no date limits for this, can have an issue if usage only increases or decreases with temperature, then it will throw an error when trying to split into both "heating" and "cooling"

weather <- read_csv("~/R/energy_split/data/recentdatedweather.csv", 
    col_types = cols(CDD = col_double(), 
        DryBulb = col_double(), HDD = col_double()))
weather$date <- as.Date(ymd(weather$WeatherDate,tz="America/New_York"))
weather <- weather %>% 
  select(-X1)

export <- inner_join(export,weather, c("date"="date"))

```

## Usage v. Temperature

Having a variable for month and weekday would be more ideal helpful, but more complicated.

```{r chart,echo=FALSE}
ggplot(export,aes(DryBulb,USAGE))+
  geom_point()+
  geom_rug(alpha=.3)+
  geom_smooth(se=F)
```

## Determine Setpoint

The setpoint is determined by the temperature at the lowest value in the above chart. The assumed setpoint is:

```{r setpoint,echo=FALSE}
p <- ggplot(export,aes(DryBulb,USAGE)) + 
      geom_smooth() #create smoothed line
model_data <- ggplot_build(p)$data[[1]][,1:5] #export x/y of smoothed line
setpoint <- model_data[order(model_data$y),][[1]][1] #get temp at lowest point
setpoint #setpoint ends up being the lowest point of the smoothed line
```

## (Re)Calculate HDD & CDD

Then, run the ol' linear model.

```{r lm,echo=FALSE}
#create HDD/CDD with new setpoint
export$HDD = ifelse(export$DryBulb > setpoint, 0, setpoint - export$DryBulb)
export$CDD = ifelse(export$DryBulb < setpoint, 0, export$DryBulb - setpoint)

#get outputs of linear model
digits <- tidy(lm(USAGE ~ HDD + CDD, export))[[2]]
datatable(tidy(lm(USAGE ~ HDD + CDD, export)), options = list(dom = 't'))
datatable(glance(lm(USAGE ~ HDD + CDD, export)), options = list(dom = 't'))
```


## Now Normalize for Typical Meterological Year 3

Based on the above setpoint, TMY 3 has these HDDs and CDDs:

```{r HDDCDD,echo=FALSE}

TMY3 <- read_csv("~/R/energy_split/data/TMY3.CSV")

#create HDD/CDD for TMY3 based on above setpoint
TMY3summary <- TMY3 %>% 
  mutate(HDH = ifelse(Fdg < setpoint,setpoint - Fdg,0),CDH = ifelse(Fdg > setpoint,Fdg - setpoint,0)) %>% 
  summarise(HDD = sum(HDH)/24,CDD = sum(CDH)/24)

TMY3summary
```

Which, given our fit above leads to the following breakout (yearly load on top, percentage of total load on the bottom):

```{r breakout,echo=FALSE}
#attempts to breakout yearly usage given model
#so takes in temperature for the day and returns what the usage should have been
usage <- data.frame(digits[1] * 365)
colnames(usage)[1] <- 'Baseload'
usage$HeatingLoad <- digits[2] * TMY3summary[[1]]
usage$CoolingLoad <- digits[3] * TMY3summary[[2]]
datatable(rbind(usage,usage/sum(usage[1:3])), options = list(dom = 't'))
```

## Usage Charts

Actual Daily Usage:

```{r Actual,echo=FALSE}

ggplot(export,aes(date,USAGE)) + 
  geom_point(alpha=.1)  + 
  geom_smooth(span=0.3,se=F)
export$Baseload <- rep(digits[1],nrow(export))
export %>% 
  mutate(HeatingLoad = HDD * digits[2],CoolingLoad = CDD * digits[3]) %>%
  gather(LoadType,Load,12:14) %>% 
  ggplot(aes(date,Load,color = LoadType)) + 
    geom_point(alpha=.3) + geom_smooth(span=0.3) + 
    theme(legend.position="bottom")

```

TMY 3 Usage:

```{r TMY32,echo=FALSE}
#what usage would have been given TMY 3 and modeled response to HDD and CDD
TMY3$hour <-( c(0, rep(1:(nrow(TMY3)-1))) + 1 )
TMY3$Baseload <- rep(digits[1]/24,nrow(TMY3))

TMY3 %>% 
  mutate(HeatingLoad = digits[2]*ifelse(Fdg < setpoint,setpoint - Fdg,0)/24,CoolingLoad = digits[3]*ifelse(Fdg > setpoint,Fdg - setpoint,0)/24,HourlyLoad = Baseload + HeatingLoad + CoolingLoad) %>%
  ggplot(aes(hour,HourlyLoad)) + 
    geom_point(alpha=.1) +
    geom_smooth()

TMY3 %>% 
  mutate(HeatingLoad = digits[2]*ifelse(Fdg < setpoint,setpoint - Fdg,0)/24,CoolingLoad = digits[3]*ifelse(Fdg > setpoint,Fdg - setpoint,0)/24) %>%
  gather(LoadType,Load,3:5) %>% 
  ggplot(aes(hour,Load,color = LoadType)) +
    geom_smooth(se=F) + 
    geom_point(alpha=.1) + 
    theme(legend.position="bottom")

```

## Actual Residuals

```{r residuals,echo=FALSE}

#how bad is this model really?
export %>% 
  mutate(HeatingLoad = HDD * digits[2],CoolingLoad = CDD * digits[3],Residual = Baseload + HeatingLoad + CoolingLoad - USAGE)  %>% 
  ggplot(aes(date,Residual,group = 1)) + 
  geom_point()+
  geom_hline(yintercept=0)


#how bad is it by day of the week
export %>% 
  mutate(HeatingLoad = HDD * digits[2],CoolingLoad = CDD * digits[3],Residual = Baseload + HeatingLoad + CoolingLoad - USAGE)  %>% 
  ggplot(aes(date,Residual,group = 1)) + 
    geom_point() + 
    facet_grid(wday(date,label = TRUE)~.)


```
